% =====================================================================
% Appendix D — Naive baseline implementations
% =====================================================================

\section{Naive baselines}
\label{app:baselines}

Three naive baselines are run on the same forecast issuance points as
the trained models, providing a floor for evaluating model
contribution. Implementations live in
\texttt{scripts/baselines/naive.py}.

\paragraph{Persistence-1d.}
$\hat y[h, z] = \mathrm{history}[-24 + h, z]$ — yesterday at the same
hour. Respects diurnal periodicity but not weekday vs weekend.

\paragraph{Persistence-7d.}
$\hat y[h, z] = \mathrm{history}[-168 + h, z]$ — same hour last week.
Respects weekly periodicity (weekday vs weekend, holiday-vs-not has its
own confound when the holiday wraps).

\paragraph{Climatological mean (4-week).}
For each forecast hour $h \in [0, 24)$ and each zone $z$, average over
all history hours whose (hour-of-day, day-of-week) tuple matches the
forecast hour's, drawn from the most-recent 4 weeks:
\begin{equation}
\hat y[h, z] = \mathrm{mean}\bigl\{\,
\mathrm{history}[i, z]\ :\
i \in \mathrm{recent-672h}\
\text{and}\
(\mathrm{hour}(i), \mathrm{dow}(i)) = (\mathrm{hour}(h), \mathrm{dow}(h))
\,\bigr\}.
\end{equation}
A 4-week window is the same as Chronos's context length. If no exact
(hour, dow) match exists in the recent window, falls back to same-hour
mean; final fallback is the zone-wise mean of the entire history.

These three baselines are universally accepted as floor comparators in
the load-forecasting literature \citep{hong2016probabilistic,
hippert2001neural} and require no training. We do not run an ARIMA /
Holt-Winters baseline in the main paper (cited in
\S\ref{sec:setup}); these three together cover the relevant signal
floor for a workshop case study.
