% =====================================================================
% Appendix E — Future-weather forecast vs analyses shift
% =====================================================================

\section{Future-weather forecast vs analyses shift}
\label{app:forecast_shift}

The trained baseline saw HRRR f00 analyses for both the history
\emph{and} future weather windows during training (privileged future
information). At deployment we cannot peek at future analyses, so we
substitute HRRR f01..f24 forecasts from the most recent long cycle
issued before $T$. This appendix isolates the input-distribution shift
this substitution introduces.

\paragraph{Procedure.}
We pick five in-training-window dates with full HRRR forecast cycle
coverage (2022-12-14, -19, -22, -28, -30) and run the baseline twice
per date:
\begin{itemize}\setlength\itemsep{1pt}
  \item Setting A (training-distribution): future weather $=$ 24 HRRR
    f00 analyses for valid hours $[T, T+23]$.
  \item Setting B (deployable):              future weather $=$ HRRR
    f01..f24 from the latest long cycle issued before $T$.
\end{itemize}
Both settings use the same history-weather (24 f00 analyses for
$[T-24, T-1]$). The only thing that differs is the future-weather
input. Implementation:
\texttt{scripts/validation/future\_weather\_shift\_quantify.py}.

\paragraph{Result.}
Across the 5 dates, the deployable setting (B) adds an MAPE penalty of
\PHFCSHIFTPP{} pp [95~\% CI: \PHFCSHIFTLO{}, \PHFCSHIFTHI{}] relative
to the training-distribution setting (A). The penalty is smaller than
the deployment-period drift documented in §\ref{sec:drift} (a few
tenths of a percentage point vs $\sim 20$ pp), so substituting
forecasts for analyses is not the dominant explanation for the
year-over-year drift.

\paragraph{Per-date breakdown.}
Specific dates and per-zone numbers are saved in
\texttt{report/arxiv/data/future\_weather\_shift.json}. The per-date
gap is small and approximately stationary across the five dates, with
the largest single-date gap on \PHFCMAXDATE{}.
