% =====================================================================
% Appendix E — Attention extraction procedure
% =====================================================================

\section{Attention extraction procedure}
\label{app:attention}

The four orientation-correctness checks used by \texttt{scripts/attention\_maps.py} (\S\ref{sec:attn}):

\begin{enumerate}\setlength\itemsep{1pt}
  \item \textbf{Slice direction}. Rows of the sliced attention matrix are future-tabular query positions $\mathcal{Q}_{\text{fut}}$; columns are history-spatial key positions $\mathcal{K}_{\text{hist}}$ (Equation~\ref{eq:slice}).
  \item \textbf{Row-major reshape}. The 64-cell column dimension is reshaped to $(8, 8)$ row-major, matching \texttt{WeatherCNN.forward}'s \texttt{x.flatten(2)} layout: cell index $i$ $\to$ (row=$i\div 8$, col=$i\bmod 8$).
  \item \textbf{Compass orientation}. \texttt{matplotlib.imshow(..., origin='upper')} so row 0 (north) is at the top, col 0 (west) is on the left.
  \item \textbf{Aggregate east-vs-west sanity}. Equation~\ref{eq:sanity}; if violated, the script emits a \texttt{[WARN]} and the figure is flagged in the report rather than aborted.
\end{enumerate}

\paragraph{The slice operation.}

\begin{equation}
A^{(\ell, h)}_{\text{slice}} \;=\; A^{(\ell, h)}\bigl[\mathcal{Q}_{\text{fut}}, \mathcal{K}_{\text{hist}}\bigr]
\;\in\; \mathbb{R}^{|\mathcal{Q}_{\text{fut}}| \times |\mathcal{K}_{\text{hist}}|}
\;=\; \mathbb{R}^{24 \times 1536},
\label{eq:slice}
\end{equation}

with $|\mathcal{Q}_{\text{fut}}| = 24$ future-tabular query positions and $|\mathcal{K}_{\text{hist}}| = S \times P = 24 \times 64 = 1536$ history-spatial key positions. We then sum over the history time axis to get a $24 \times 64$ map of future-time vs spatial position, and reshape the spatial axis to $8 \times 8$.

\paragraph{The east-vs-west sanity check.}

\begin{equation}
\sum_{r=0}^{7} \sum_{c=4}^{7} A_{\text{agg}}[r, c]
\;>\;
\sum_{r=0}^{7} \sum_{c=0}^{3} A_{\text{agg}}[r, c],
\label{eq:sanity}
\end{equation}

i.e., aggregate attention mass on the eastern (cols 4--7) half of the bbox should exceed the western half because dense load (Boston metro, southern Massachusetts) sits in the eastern half. Observed values: east $\approx 0.0070$, west $\approx 0.0050$ --- check passes.

\paragraph{Diagnostic days.} The five hand-picked days used in \S\ref{sec:attn} are 2022-04-15 (mild), 2022-07-21 (heat-wave), 2022-08-04 (extreme), 2022-12-23 (cold-snap), and 2022-12-25 (holiday). Drawn from the broader 2022 calendar; the held-out test slice 2022-12-30/31 is too narrow to span all five regimes.
