% =====================================================================
% Appendix B — ISO-NE cookie-prime endpoint
% =====================================================================

\section{ISO-NE cookie-prime endpoint and zone-ID mapping}
\label{app:isone}

ISO New England exposes a public 5-minute estimated zonal load CSV at
\begin{quote}\small\texttt{https://www.iso-ne.com/transform/csv/fiveminuteestimatedzonalload?start=YYYYMMDD\&end=YYYYMMDD}\end{quote}
Direct \texttt{curl} to this endpoint returns HTTP 403. The endpoint requires a session cookie that is set when a browser visits any normal ISO-NE report page (e.g.\ the gen-fuel-mix tree). The deployed fetcher (\texttt{space/iso\_ne\_zonal.py}) primes that cookie at session start and reuses it for all subsequent CSV requests:

\begin{lstlisting}[language=Python, caption={Cookie-prime trick (extract).}, label={lst:isone}]
import requests
session = requests.Session()
session.get(
    "https://www.iso-ne.com/isoexpress/web/reports/operations/-/tree/gen-fuel-mix"
)  # primes session cookie
url = (
    "https://www.iso-ne.com/transform/csv/fiveminuteestimatedzonalload"
    f"?start={yyyymmdd}&end={yyyymmdd}"
)
r = session.get(url, timeout=15)
\end{lstlisting}

\paragraph{Zone-ID mapping.} The CSV uses ISO-NE's internal numeric zone IDs, which we map to the canonical column names:

\begin{center}\small
\begin{tabular}{@{}rl@{\hspace{1.5em}}rl@{}}
\toprule
ID & Column & ID & Column \\
\midrule
4001 & \zone{ME}   & 4005 & \zone{RI}        \\
4002 & \zone{NH}   & 4006 & \zone{SEMA}      \\
4003 & \zone{VT}   & 4007 & \zone{WCMA}      \\
4004 & \zone{CT}   & 4008 & \zone{NEMA\_BOST}\\
\bottomrule
\end{tabular}
\end{center}

\paragraph{5-min $\to$ hourly aggregation.} The native CSV has 12 5-minute samples per hour per zone. We resample to hourly mean to match the model's input shape \texttt{(24, 8)}. Publication lag is typically $\sim$1--2 hours: the most recent fully-published hour at request time is approximately \texttt{now() - 2h}.

\paragraph{Why this trick works.} The cookie-prime pattern is borrowed from the open-source \texttt{gridstatus.io} library, which uses the same approach for several ISO-NE endpoints. The cookie is short-lived (session duration only); we re-prime on each Python process start. There is no rate-limit on the documented endpoint as of writing; we have observed no 429s in production.
